Calculating Entropies With Nested Sampling

### 2017-02-04 paper/ms.tex
Changed x to theta, and also wrote in a few more relations
Updated example to make it more computable

\renewcommand{\d}{\boldsymbol{d}}
\newcommand{\todo}{\color{orange} \bf}
\newcommand{\query}{\color{green} \bf}
\newcommand{\x}{\boldsymbol{x}}

@@

In a standard situation where Bayesian inference would be used,
the unknown quantities are the parameters $\x$ and the data $\d$
(which subsequently becomes known). The prior information leads the
user to specify a prior distribution $p(\x)$ for the unknown parameters,
along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
of how the data is related to the parameters. By the product rule, this yields
a {\em joint prior}
\begin{align}
p(\x, \d) &= p(\x)p(\d | \x)
\end{align}

@@

The degree to which the question ``what is the value of $\x$?'' is
answered by the question ``what is the value of $\d$?'' is given by the
conditional entropy
\begin{align}
H_{\x | \d} &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
\end{align}
which is the expected value of the entropy of the posterior, averaged over
all possible datasets which might be observed.

This is difficult to compute because the expression
for the posterior distribution
\begin{align}
p(\x | \d) &= \frac{p(\x)p(\d | \x)}{p(\d)}
\end{align}
contains the marginal likelihood $p(\d)$.

@@

\subsection{Mutual information}
The mutual information is another way of describing the relevance of the
data to the parameters.

Its definition is
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
                       \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
           &= H_{\x} + H_{\d} - H_{\x, \d}.
\end{align}
For present purposes, it is more conveniently written as
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
              \log\left[\frac{p(\d | \x)}{p(\d)}\right].
\end{align}
As with the conditional entropy, we see that the difficulty appears in the
form of the log of the marginal likelihood,
$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.

@@

The priors are
\begin{align}
A           &\sim \textnormal{Exponential}(1)  \\
\log_{10} T &\sim \textnormal{Uniform}(-3, 0)  \\
\phi        &\sim \textnormal{Uniform}(0, 2\pi)
\end{align}
$A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
These results were computed using a tolerance of 0.001 for every quantity.
Results:

@@

\begin{align}                    % (num_particles, mcmc_steps, max_depth, reps)
H_{\x}    &= 7.328 \pm 0.025 \\% (100, 1000, 30.0,  122)
H_{\d}    &= 960.4 \pm 3.7   \\% (100, 1000, 1200.0, 99)
H_{\x,\d} &= 968.6 \pm 3.4   \\% (100, 1000, 1200.0, 99)
\end{align}

>>>>

\renewcommand{\d}{\boldsymbol{d}}
\newcommand{\todo}{\color{orange} \bf}
\newcommand{\query}{\color{green} \bf}
\newcommand{\x}{\boldsymbol{\theta}}

@@

Bayesian inference is the use of probability theory to
describe uncertainty, often about unknown quantities
(``parameters'') $\x$. Some data $\d$, initially unknown but
thought to be relevant to $\x$, is obtained.
The prior information leads the
user to specify a prior distribution $p(\x)$ for the unknown parameters,
along with a ``sampling distribution'' $p(\d | \x)$ describing prior knowledge
of how the data is related to the parameters
(i.e., if the parameters where known, what data would likely be observed?). By the product rule, this yields a {\em joint prior}
\begin{align}
p(\x, \d) &= p(\x)p(\d | \x)
\end{align}
which is the starting point for inference.

@@

The degree to which the question ``what is the value of $\x$?'' is
answered by the question ``what is the value of $\d$?'' is given by the
conditional entropy
\begin{align}
H_{\x | \d} &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
\end{align}
which is the expected value of the entropy of the posterior, averaged over
all possible datasets which might be observed.
This is difficult to compute because the expression
for the posterior distribution
\begin{align}
p(\x | \d) &= \frac{p(\x)p(\d | \x)}{p(\d)}
\end{align}
contains the marginal likelihood $p(\d)$.

@@

\subsection{Mutual information}
The mutual information is another way of describing the relevance of the
data to the parameters. Its definition, and relation to other quantities, is
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
                       \log\left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]\\
           &= H_{\x} + H_{\d} - H_{\x, \d}\\
           &= H_{\d} - H_{\d | \x}\\
           &= H_{\x} - H_{\x | \d}.
\end{align}
For present purposes, it is more conveniently written as
\begin{align}
I_{\x, \d} &= \sum_{\x} \sum_{\d} p(\x, \d)
              \log\left[\frac{p(\d | \x)}{p(\d)}\right].
\end{align}
As with the conditional entropy, we see that the difficulty appears in the
form of the log of the marginal likelihood,
$\log p(\d) = \log \sum_{\x} p(\x)p(\d | \x)$.

@@

The priors are
\begin{align}
A           &\sim \textnormal{Exponential}(1)  \\
\log_{10} T &\sim \textnormal{Uniform}(-1, 0)  \\
\phi        &\sim \textnormal{Uniform}(0, 2\pi)
\end{align}
$A$ and $\phi$ are nuisance parameters. Let $\x = \log_{10} T$.
The prior standard deviation of the noise, $\sigma$, was set to 0.1.
These results were computed using a tolerance of 0.001 for every quantity.
Results:

@@

\begin{align}                    % (num_particles, mcmc_steps, max_depth, reps)
H_{\x}    &= 6.223 \pm 0.021 \\% (100, 1000, 20.0,  147)
H_{\d}    &=    \\% (100, 1000, 1200.0, 99)
H_{\x,\d} &=    \\% (100, 1000, 1200.0, 99)
\end{align}
